---
title: "Lowland Mineral Abrupt Thaw Model using FME as reported by Turetsky et al. (2019) Nature Biogeosciences"
author: "Merritt Turetsky"
date: "June 15, 2019"
output: pdf_document
editor_options:
  chunk_output_type: console

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE,
                      fig.align = "center",
                      collapse = TRUE, comment = "#>")
```

```{r libraries}
library(tidyverse)
library(FME)
library(stringr)
```

# Variable Names
- *undisturbed_yedoma* - permafrost tundra in yedoma soils
- *young_lake_yedoma* - active thermokarst lake features that persist for 250 yrs before transitioning into stabilized thaw lakes inin yedoma soil
- *old_lake_yedoma* -  stabilized thermokarst lake features that persist for 3000 yrs in yedoma soil before transitioning to drained lake basins
- *alas_yedoma* - drained lake basins that persist for 3000 years in yedoma soil before transitioning to permafrost tundra
- *undisturbed_non* - permafrost tundra in non-yedoma soils
- *young_lake_non* - active thaw lake features that persist for 250 yrs before transitioning into stabilized thaw lakes in non-yedoma soil
- *old_lake_non* -  stabilized thermokarst lake features that persist for 3000 yrs in non-yedoma soil before transitioning to drained lake basins
- *alas_non* - drained lake basins that persist for 3000 years in non-yedoma soil before transitioning to permafrost tundra
- *undisturbed_non_0* - starting undisturbed permafrost area of yedoma mineral terrain
- *young_lake_non_0* - starting non-yedoma young/active thermokarst lake area
- *old_lake_non_0* - starting non-yedoma old/stabilized thermokarst lake area 
- *alas__non_0* - starting non-yedoma alas/drained lake basin area
- *undisturbed_yedoma_0* - starting undisturbed area of yedoma mineral terrain
- *young_lake__yedoma0* - starting young/active thermokarst fen area in lake model
- *old_lake__yedoma0* - starting old/stabilized thermokarst lake area in lake model
- *alas__yedoma_0* - starting alas/drained lake basin area in lake model
- *CH4* - Methane emissions (g C-CH4 m-2 yr-1)
- *NEE* - Net ecosystem exchange (g C-CO2 m-2 yr-1)
- *DOC* - Dissolved organic carbon export flux (g C m-2 yr-1)

Transition rates are the same for yedoma and nonyedoma lakes
- *a* - rate at which permafrost tundra is converted to young/active thermokarst lake
- *b* - rate at which young/active thermokrast lake is converted to mature/stabilized thermokarst lake
- *c* - rate at which mature thermokarst lake is converted to alas/drained lake basin
- *d* - rate at which alas is converted to permafrost tundra

Starting areas are defined within the function.

Note that *a* changes through time by increasing by an amount *a_dynamic* which is set as an input parameter. *d* also changes through time to reflect increasing landscape drying and loss of limnicity by an amount *d_dynamic*.  *b* and *c* are fixed and are entered as input paramters to the function.

#Lake model
##The order that these parameters are introduced is how subsequent lines of code will be calculated. The name of the variable doesn't imply a link to that variable in subsequent code (names don't matter). The order matters.
```{r function}
NECB_model <- function(pars, start_time, end_time) {
  derivs <- function(time, y, pars) {
    with(as.list(c(pars, y)), {
      dundisturbed_yedoma <- -a_yedoma * undisturbed_yedoma + d_yedoma * alas_yedoma
      
      dyoung_lake_yedoma <- a_yedoma * undisturbed_yedoma - b_yedoma * young_lake_yedoma
      
      dold_lake_yedoma <- b_yedoma * young_lake_yedoma - c_yedoma * old_lake_yedoma
      dalas_yedoma <- c_yedoma * old_lake_yedoma - d_yedoma * alas_yedoma
      da_yedoma <- a_dynamic_yedoma
      dc_yedoma <- c_dynamic_yedoma
      dundisturbed_yedoma_NECB <- ((undisturbed_yedoma * 1e6 * (undisturbed_yedoma_NEE + undisturbed_yedoma_CH4 + undisturbed_yedoma_DOC))) / 1e15
      dyoung_lake_yedoma_NECB <- ((young_lake_yedoma * 1e6 * (young_lake_yedoma_NEE + young_lake_yedoma_CH4 + young_lake_yedoma_DOC))) / 1e15
      dold_lake_yedoma_NECB <- ((old_lake_yedoma * 1e6 * (old_lake_yedoma_NEE + old_lake_yedoma_CH4 + old_lake_yedoma_DOC))) / 1e15
      dalas_yedoma_NECB <- ((alas_yedoma * 1e6 * (alas_yedoma_NEE + alas_yedoma_CH4 + alas_yedoma_DOC))) / 1e15
      dTotal_yedoma_NECB <- dundisturbed_yedoma_NECB + dyoung_lake_yedoma_NECB + dold_lake_yedoma_NECB + dalas_yedoma_NECB
      
      dundisturbed_non <- -a_non * undisturbed_non + d_non * alas_non
      dyoung_lake_non <- a_non * undisturbed_non - b_non * young_lake_non
      dold_lake_non <- b_non * young_lake_non - c_non * old_lake_non
      dalas_non <- c_non * old_lake_non - d_non * alas_non
      da_non <- a_dynamic_non
      dc_non <- c_dynamic_non
      dundisturbed_non_NECB <- ((undisturbed_non * 1e6 * (undisturbed_non_NEE + undisturbed_non_CH4 + undisturbed_non_DOC))) / 1e15
      dyoung_lake_non_NECB <- ((young_lake_non * 1e6 * (young_lake_non_NEE + young_lake_non_CH4 + young_lake_non_DOC))) / 1e15
      dold_lake_non_NECB <- ((old_lake_non * 1e6 * (old_lake_non_NEE + old_lake_non_CH4 + old_lake_non_DOC))) / 1e15
      dalas_non_NECB <- ((alas_non * 1e6 * (alas_non_NEE + alas_non_CH4 + alas_non_DOC))) / 1e15
      dTotal_non_NECB <- dundisturbed_non_NECB + dyoung_lake_non_NECB + dold_lake_non_NECB + dalas_non_NECB
      dundisturbed_total <- dundisturbed_non + dundisturbed_yedoma
      dyoung_lake_total <- dyoung_lake_non + dyoung_lake_yedoma
      dold_lake_total <- dold_lake_non + dold_lake_yedoma
      dalas_total <- dalas_non + dalas_yedoma

      dCH4 <- (((undisturbed_non * 1e6 * (undisturbed_non_CH4))) / 1e15) + (((young_lake_non * 1e6 * (young_lake_non_CH4))) / 1e15) + (((old_lake_non * 1e6 * (old_lake_non_CH4))) / 1e15) + (((alas_non * 1e6 * (alas_non_CH4))) / 1e15) + (((undisturbed_yedoma * 1e6 * (undisturbed_yedoma_CH4))) / 1e15) + (((young_lake_yedoma * 1e6 * (young_lake_yedoma_CH4))) / 1e15) + (((old_lake_yedoma * 1e6 * (old_lake_yedoma_CH4))) / 1e15) + (((alas_yedoma * 1e6 * (alas_yedoma_CH4))) / 1e15)
      
      dCO2 <- (((undisturbed_non * 1e6 * (undisturbed_non_NEE))) / 1e15) + (((young_lake_non * 1e6 * (young_lake_non_NEE))) / 1e15) + (((old_lake_non * 1e6 * (old_lake_non_NEE))) / 1e15) + (((alas_non * 1e6 * (alas_non_NEE))) / 1e15) + (((undisturbed_yedoma * 1e6 * (undisturbed_yedoma_NEE))) / 1e15) + (((young_lake_yedoma * 1e6 * (young_lake_yedoma_NEE))) / 1e15) + (((old_lake_yedoma * 1e6 * (old_lake_yedoma_NEE))) / 1e15) + (((alas_yedoma * 1e6 * (alas_yedoma_NEE))) / 1e15)
      
      dTotal_NECB <- dTotal_yedoma_NECB + dTotal_non_NECB

      return(list(c(
        dundisturbed_yedoma,
        dyoung_lake_yedoma,
        dold_lake_yedoma,
        dalas_yedoma,
        da_yedoma,
        dc_yedoma,
        dundisturbed_yedoma_NECB,
        dyoung_lake_yedoma_NECB,
        dold_lake_yedoma_NECB,
        dalas_yedoma_NECB,
        dTotal_yedoma_NECB,
        dundisturbed_non,
        dyoung_lake_non,
        dold_lake_non,
        dalas_non,
        da_non,
        dc_non,
        dundisturbed_non_NECB,
        dyoung_lake_non_NECB,
        dold_lake_non_NECB,
        dalas_non_NECB,
        dTotal_non_NECB,
        dundisturbed_total,
        dyoung_lake_total,
        dold_lake_total,
        dalas_total,
        dCH4,
        dCO2,
        dTotal_NECB)))
    })
  }
  # initial conditions
  y <- with(as.list(pars), {
    c(
      undisturbed_yedoma = undisturbed_yedoma_0,
      young_lake_yedoma = young_lake_yedoma_0,
      old_lake_yedoma = old_lake_yedoma_0,
      alas_yedoma = alas_yedoma_0,
      a_yedoma = a_yedoma_0,
      c_yedoma = c_yedoma_0,
      undisturbed_yedoma_NECB = undisturbed_yedoma_NECB_0,
      young_lake_yedoma_NECB = young_lake_yedoma_NECB_0,
      old_lake_yedoma_NECB = old_lake_yedoma_NECB_0,
      alas_yedoma_NECB = alas_yedoma_NECB_0,
      Total_yedoma_NECB = Total_yedoma_NECB_0,
      undisturbed_non = undisturbed_non_0,
      young_lake_non = young_lake_non_0,
      old_lake_non = old_lake_non_0,
      alas_non = alas_non_0,
      a_non = a_non_0,
      c_non = c_non_0,
      undisturbed_non_NECB = undisturbed_non_NECB_0,
      young_lake_non_NECB = young_lake_non_NECB_0,
      old_lake_non_NECB = old_lake_non_NECB_0,
      alas_non_NECB = alas_non_NECB_0,
      Total_non_NECB = Total_non_NECB_0,
      undisturbed_total = undisturbed_total_0,
      young_lake_total = young_lake_total_0,
      old_lake_total = old_lake_total_0,
      alas_total = alas_total_0,
      CH4 = CH4_0,
      CO2 = CO2_0,
      Total_NECB = Total_NECB_0)
  })
  times <- start_time:end_time
  out <- ode(y = y, parms = pars, times = times, func = derivs)
  as.data.frame(out)
}
```

# Parameter Values for Static/Historical Lake Model 
##This code is used as an informal means of model spin up, mainly so that CO2 and CH4 release from these thaw landscapes do not start at 0 at the beginning of our dynamic measurement period.  In the results reported in Turetsky et al. (2019) we substract carbon emissions from 1900-2000 from our results and focus mainly on emissions from 2000-2300.  Parameters in the historical model run were fitted so that abrupt thaw rates approximated an equilibrium with permafrost recovery at the regional scale. This allows us to explore net ecosystem carbon balance at regional scales under these equilibrium conditions, and then compare changes in net ecosystem carbon balance under accelerated thaw rates in the dynamic measurement period.
##Dynamic increases in transition rates are set to zero in this simulation
```{r values for historical measurement period}
lake_pars_historical <- c(
  a_yedoma_0 = 0.0002,
  a_dynamic_yedoma = 0,
  b_yedoma = 0.01,
  c_yedoma_0 = 0.001,
  c_dynamic_yedoma = 0,
  d_yedoma = 0.0003, 
  undisturbed_yedoma_0 = 171750, 
  young_lake_yedoma_0 = 2370, 
  old_lake_yedoma_0 = 76630, 
  alas_yedoma_0 = 57250, 
  undisturbed_yedoma_NECB_0 = 0,
  young_lake_yedoma_NECB_0 = 0,
  old_lake_yedoma_NECB_0 = 0,
  alas_yedoma_NECB_0 = 0,
  Total_yedoma_NECB_0 = 0,
  undisturbed_yedoma_NEE = 10.5,
  undisturbed_yedoma_CH4 = -5,
  undisturbed_yedoma_DOC = -2.5,
  young_lake_yedoma_NEE = -450,  
  young_lake_yedoma_CH4 = -130,  
  young_lake_yedoma_DOC = 0,
  old_lake_yedoma_NEE = -181, 
  old_lake_yedoma_CH4 = -10, 
  old_lake_yedoma_DOC = 0,
  alas_yedoma_NEE = 26,
  alas_yedoma_CH4 = -5,
  alas_yedoma_DOC = 0,
  b_non = 0.01,
  c_non_0 = 0.001,
  c_dynamic_non = 0,
  d_non = 0.0003, 
  undisturbed_non_0 = 602550, 
  young_lake_non_0 = 3618, 
  old_lake_non_0 = 116982, 
  alas_non_0 = 200850,
  a_non_0 = 0.0002,
  a_dynamic_non = 0,
  undisturbed_non_NECB_0 = 0,
  young_lake_non_NECB_0 = 0,
  old_lake_non_NECB_0 = 0,
  alas_non_NECB_0 = 0,
  Total_non_NECB_0 = 0,
  undisturbed_non_NEE = 10.5,
  undisturbed_non_CH4 = -5,
  undisturbed_non_DOC = -3,
  young_lake_non_NEE = -149,  
  young_lake_non_CH4 = -38,  
  young_lake_non_DOC = 0,
  old_lake_non_NEE = -54,  
  old_lake_non_CH4 = -7,
  old_lake_non_DOC = 0,
  alas_non_NEE = 30, 
  alas_non_CH4 = -3,
  alas_non_DOC = 0,
  undisturbed_total_0 = 774300,
  young_lake_total_0 = 5988,
  old_lake_total_0 = 193612,
  alas_total_0 = 258100,
  CH4_0 = 0,
  CO2_0 = 0,
  Total_NECB_0 = 0
)
```

# Run Static/historical Lake Model
##End point = initial conditions for next model
```{r run historical measurement period}
lake_out_historical <- NECB_model(pars = lake_pars_historical, start_time = 1900, end_time = 2000)
end_values<-subset(lake_out_historical, lake_out_historical$time==2000)
```

# Parameter Values for Dynamic Lake Model 
##This includes a dynamic increase in thaw rate to simulate the effects of climate change on permafrost area. The dynamic increase in thaw rate is a_dynamic, which was fitted so that the cumulative change in permafrost area by 2300 matched that of large scale modeling of gradual thaw reported in McGuire et al. 2018.
##In Turetsky et al. (2019), the RCP8.5 scenario used an *a_dynamic* of 0.00003 for yedoma and 0.000013 for nonyedoma while the RCP4.5 scenario used an *a_dynamic* of 0 for yedoma and nonyedoma meaning there was no dynamic increase in thaw rate over time. Rather for the RCP4.5 scenario, a total reduction in permafrost extent tracking that of large scale gradual thaw models was achieved using the static thaw rate inputted at the start of the model run.
##This code pulls starting values from the end of the 'historical' run of the model.
```{r values for dynamic measurement period}
lake_pars <- c(
  #start of yedoma parameters
  a_yedoma_0 = 0.003333333, 
  a_dynamic_yedoma = 0.00003, 
  b_yedoma = 0.004,
  c_yedoma_0 = 0.000333333,
  c_dynamic_yedoma = 0.00002, 
  d_yedoma = 0.000333333,
  undisturbed_yedoma_0 = end_values$undisturbed_yedoma,
  young_lake_yedoma_0 = end_values$young_lake_yedoma,
  old_lake_yedoma_0 = end_values$old_lake_yedoma,
  alas_yedoma_0 = end_values$alas_yedoma,
  undisturbed_yedoma_NECB_0 = end_values$undisturbed_yedoma_NECB,
  young_lake_yedoma_NECB_0 = end_values$young_lake_yedoma_NECB,
  old_lake_yedoma_NECB_0 = end_values$old_lake_yedoma_NECB,
  alas_yedoma_NECB_0 = end_values$alas_yedoma_NECB,
  Total_yedoma_NECB_0 = end_values$Total_yedoma_NECB,
  undisturbed_yedoma_NEE = 10.5,
  undisturbed_yedoma_CH4 = -5,
  undisturbed_yedoma_DOC = -3,
  young_lake_yedoma_NEE = -450, 
  young_lake_yedoma_CH4 = -130, 
  young_lake_yedoma_DOC = 0,
  old_lake_yedoma_NEE = -181, 
  old_lake_yedoma_CH4 = -10,
  old_lake_yedoma_DOC = 0,
  alas_yedoma_NEE = 26.1, 
  alas_yedoma_CH4 = -5,
  alas_yedoma_DOC = 0,
  #start of non-yedoma parameters
  a_non_0 = 0.003333333,
  a_dynamic_non = 0.000013,
  b_non = 0.004,
  c_non_0 = 0.000333333,
  c_dynamic_non = 0.00002,
  d_non = 0.000333333,
  undisturbed_non_0 = end_values$undisturbed_non,
  young_lake_non_0 = end_values$young_lake_non,
  old_lake_non_0 = end_values$old_lake_non,
  alas_non_0 = end_values$alas_non,
  undisturbed_non_NECB_0 = end_values$undisturbed_non_NECB,
  young_lake_non_NECB_0 = end_values$young_lake_non_NECB,
  old_lake_non_NECB_0 = end_values$old_lake_non_NECB,
  alas_non_NECB_0 = end_values$alas_non_NECB,
  Total_non_NECB_0 = end_values$Total_non_NECB,
  undisturbed_non_NEE = 10.5,
  undisturbed_non_CH4 = -4,
  undisturbed_non_DOC = -2,
  young_lake_non_NEE = -149, 
  young_lake_non_CH4 = -38,
  young_lake_non_DOC = 0,
  old_lake_non_NEE = -54,
  old_lake_non_CH4 = -7,
  old_lake_non_DOC = 0,
  alas_non_NEE = 29.5,
  alas_non_CH4 = -3,
  alas_non_DOC = 0,
  undisturbed_total_0 = end_values$undisturbed_total,
  young_lake_total_0 = end_values$young_lake_total,
  old_lake_total_0 = end_values$old_lake_total,
  alas_total_0 = end_values$alas_total,
  CH4_0 = end_values$CH4,
  CO2_0 = end_values$CO2,
  Total_NECB_0 = end_values$Total_NECB
)
```

# Run Dynamic Lake Model
```{r run dynamic measurement period}
lake_out_dynamic <- NECB_model(pars = lake_pars, start_time = 2000, end_time = 2300)
lake_out<-rbind(lake_out_historical %>% filter(time < 2000), 
                lake_out_dynamic)
```

#Plot Model Results
##transform the data into a more practical format
##rename old to mature
```{r model-ggplot2}
lake_model_tidy <- lake_out %>% 
  gather(variable, value, -time) %>% 
  # rename Und to Undisturbed
  mutate(
    variable_name = variable,
    variable_name = recode(variable_name,
                           undisturbed_yedoma = "Undisturbed tundra", 
                           young_lake_yedoma = "Active thaw lake", 
                           old_lake_yedoma = "Mature thaw lake", 
                           alas_yedoma = "Alas",
                           undisturbed_non = "Undisturbed tundra", 
                           young_lake_non = "Active thaw lake", 
                           old_lake_non = "Mature thaw lake", 
                           alas_non = "Alas",
                           undisturbed_total = "Undisturbed tundra", 
                           young_lake_total = "Active thaw lake", 
                           old_lake_total = "Mature thaw lake", 
                           alas_total = "Alas"))
write_csv(lake_model_tidy, "lake_model_tidy.csv")
```

#Plot Results of Changing Areas in Lake Model
```{r plot changing areas}
# yedoma area
plot_vars <- c(undisturbed_yedoma = "Undisturbed tundra", 
               young_lake_yedoma = "Active thaw lake", 
               old_lake_yedoma = "Mature thaw lake", 
               alas_yedoma = "Alas/Drained lake")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% names(plot_vars)) %>% 
  mutate(variable_name = factor(variable_name, levels = plot_vars))

ggplot(plot_data, aes(x = time, y = value, color = variable_name)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  scale_color_brewer(NULL, palette = "Set1") +
  scale_y_continuous(label = scales::comma) +
  labs(x = "Year", y = bquote("Area of Yedoma Terrain"~(km^2))) +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

# Non-yedoma area
plot_vars <- c(undisturbed_non = "Undisturbed tundra", 
               young_lake_non = "Active thaw lake", 
               old_lake_non = "Mature thaw lake", 
               alas_non = "Alas")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% names(plot_vars)) %>% 
  mutate(variable_name = factor(variable_name, levels = plot_vars))
ggplot(plot_data, aes(x = time, y = value, color = variable_name)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  scale_color_brewer(NULL, palette = "Set1") +
  scale_y_continuous(label = scales::comma) +
  labs(x = "Year", y = bquote("Area of Non-Yedoma Terrain"~(km^2))) +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())


# combined area
plot_vars <- c(undisturbed_total = "Undisturbed tundra", 
               young_lake_total = "Active thaw lake", 
               old_lake_total = "Mature thaw lake", 
               alas_total = "Alas")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% names(plot_vars)) %>% 
  mutate(variable_name = factor(variable_name, levels = plot_vars))
lake_area_fig<-ggplot(plot_data, aes(x = time, y = value, color = variable_name)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  scale_color_brewer(NULL, palette = "Set1") +
  scale_y_continuous(label = scales::comma) +
  labs(x = "Year", y = bquote("Mineral Lowland Area"~(km^2))) +
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=22),
        legend.text = element_text(size=18),
        #legend.position = "bottom", 
        legend.position= c(0.8, 0.83),
        plot.margin = margin(.8, .8, .8, .8, "cm"),
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

print(lake_area_fig)

pdf(file="lake_area_fig.pdf", width=9, height=6, family="Helvetica")
lake_area_fig
dev.off()

```

#Plot Results of Model NECB
```{r plot NECB results}
# yedoma
plot_vars <- c("undisturbed_yedoma_NECB", "young_lake_yedoma_NECB", 
               "old_lake_yedoma_NECB", "alas_yedoma_NECB", 
               "Total_yedoma_NECB")
plot_lbls <- c("Permafrost", "Active thaw lake", "Mature thaw lake", "Alas", "Total")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% plot_vars) %>% 
  mutate(variable = factor(variable, levels = plot_vars))
ggplot(plot_data, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  geom_hline(yintercept = 0) +
  scale_color_brewer(NULL, palette = "Set1", breaks = plot_vars, 
                     labels = plot_lbls) +
  labs(title = paste("Lake thermokarst Net Ecosystem Carbon Balance (NECB)", 
                     "from Yedoma Lowland"),
       subtitle = "Predicted changes in NECB",
       x = "Year", y = "Cumulative Yedoma NECB (PgC)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())

# Non-yedoma
plot_vars <- c("undisturbed_non_NECB", "young_lake_non_NECB", 
               "old_lake_non_NECB", "alas_non_NECB", 
               "Total_non_NECB")
plot_lbls <- c("Permafrost", "Active thaw lake", "Mature thaw lake", "Alas", "Total")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% plot_vars) %>% 
  mutate(variable = factor(variable, levels = plot_vars))
ggplot(plot_data, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  geom_hline(yintercept = 0) +
  scale_color_brewer(NULL, palette = "Set1", breaks = plot_vars, 
                     labels = plot_lbls) +
  labs(x = "Year", y = "Cumulative Non-Yedoma Net Ecosystem Carbon Balance (Pg C)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())


# combined
plot_vars <- c("Total_NECB", "Total_yedoma_NECB", "Total_non_NECB")
plot_lbls <- c("Total", "Yedoma", "Non-Yedoma")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% plot_vars) %>% 
  mutate(variable = factor(variable, levels = plot_vars))
ggplot(plot_data, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  geom_hline(yintercept = 0) +
  scale_color_brewer(NULL, palette = "Set1", breaks = plot_vars, 
                     labels = plot_lbls) +
  labs(x = "Year", y = "Cumulative Net Ecosystem Carbon Balance (Pg C)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```


#Plot Results of CO2 and CH4 from the entire terrain
```{r plot cumulative CO2 and CH4 emissions}
# combined
plot_vars <- c("CO2", "CH4", "CH4_CO2equiv", "Total_NECB")
plot_lbls <- c("CO2", "CH4", "CH4_CO2equiv", "NECB")
plot_data <- lake_model_tidy %>% 
  filter(variable %in% plot_vars) %>% 
  mutate(variable = factor(variable, levels = plot_vars))
ggplot(plot_data, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1.5, linetype = "dashed") +
  geom_line(data = filter(plot_data, time >= 2000), size = 1.5) +
  geom_hline(yintercept = 0) +
  scale_color_brewer(NULL, palette = "Set1", breaks = plot_vars,
                     labels = plot_lbls) +
  labs(x = "Year", y = "Cumulative Lowland Mineral Abrupt Thaw CO2 or CH4 (PgC)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```

#Local Sensistivity Lake Model
##Run a local sensitivity analysis using a subset of parameters for total NECB. After running the sensitivity analysis, transform the data to a format more conducive to analysis and plotting.
##This model has been split into yedoma and non-yedoma components. Here we look at the sensitive of the total NECB variable to changes in yedoma/non-yedoma parameters.

```{r local sensitivity NECB}
## Yedoma
sp <- c("a_yedoma_0", "a_dynamic_yedoma", "c_dynamic_yedoma", "b_yedoma", "c_yedoma_0", 
        "undisturbed_yedoma_0", "young_lake_yedoma_0", "old_lake_yedoma_0", 
        "alas_yedoma_0", "undisturbed_yedoma_NEE", "undisturbed_yedoma_CH4", 
        "undisturbed_yedoma_DOC", "old_lake_yedoma_NEE", "old_lake_yedoma_CH4", 
        "old_lake_yedoma_DOC", "alas_yedoma_NEE", "alas_yedoma_CH4", 
        "alas_yedoma_DOC")
SnsThermokarst_y <- sensFun(func = NECB_model, parms = lake_pars,
                            start_time=1900, end_time=2300, 
                            senspar = sp, sensvar = "Total_NECB", varscale = -1)
# transform
yedoma_loc_sens <- SnsThermokarst_y %>% 
  gather(parameter, sensitivity, -x, -var) %>% 
  rename(year = x, variable = var)
write_csv(yedoma_loc_sens, "lake_local-sensitivity_yedoma.csv")
```

#Average yedoma sensitivity results over years to produce a bar plot.
```{r glbl-yr-avg-yedoma}
sens_summ_yedoma <- bind_rows(
  yedoma_loc_sens %>% 
    filter(variable == "Total_NECB", year >= 2000, year <= 2100) %>% 
    group_by(parameter) %>% 
    summarize(sens_mean = mean(sensitivity),
              sens_se = sd(sensitivity) / n()) %>% 
    mutate(period = "2000-2100"),
  yedoma_loc_sens %>% 
    filter(variable == "Total_NECB", year >= 2000, year <= 2300) %>% 
    group_by(parameter) %>% 
    summarize(sens_mean = mean(sensitivity),
              sens_se = sd(sensitivity) / n()) %>% 
    mutate(period = "2000-2300")
)
ggplot(sens_summ_yedoma, aes(x = parameter, y = sens_mean)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~ period) +
  labs(x = "Sensitivity", y = "Parameter",
       title = "Mean local sensitivity for lake thermokarst NECB (Yedoma)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```


```{r}
# sensitivity parameters nonyedoma
sp <- c("a_non_0", "a_dynamic_non", "b_non", #"c_non", 
        "undisturbed_non_0", "young_lake_non_0", "old_lake_non_0", 
        "alas_non_0", "undisturbed_non_NEE", "undisturbed_non_CH4", 
        "undisturbed_non_DOC", "old_lake_non_NEE", "old_lake_non_CH4", 
        "old_lake_non_DOC", "alas_non_NEE", "alas_non_CH4", 
        "alas_non_DOC")
SnsThermokarst_n <- sensFun(func = NECB_model, parms = lake_pars, 
                            start_time=1900, end_time=2300, senspar = sp, 
                            sensvar = "Total_NECB", varscale = -1)
# transform
non_loc_sens <- SnsThermokarst_n %>% 
  gather(parameter, sensitivity, -x, -var) %>% 
  rename(year = x, variable = var)
write_csv(non_loc_sens, "lake_local-sensitivity_non-yedoma.csv")
```

#Average over years to produce a bar plot.
```{r glbl-yr-avg-non}
sens_summ_non <- bind_rows(
  non_loc_sens %>% 
    filter(variable == "Total_NECB", year >= 2000, year <= 2100) %>% 
    group_by(parameter) %>% 
    summarize(sens_mean = mean(sensitivity),
              sens_se = sd(sensitivity) / n()) %>% 
    mutate(period = "2000-2100"),
  non_loc_sens %>% 
    filter(variable == "Total_NECB", year >= 2000, year <= 2300) %>% 
    group_by(parameter) %>% 
    summarize(sens_mean = mean(sensitivity),
              sens_se = sd(sensitivity) / n()) %>% 
    mutate(period = "2000-2300")
)
ggplot(sens_summ_non, aes(x = parameter, y = sens_mean)) +
  geom_point() +
  coord_flip() +
  facet_wrap(~ period) +
  labs(x = "Sensitivity", y = "Parameter",
       title = "Mean local sensitivity for lake thermokarst NECB (Non-Yedoma)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```

# Global Sensistivity Lake Model
##This analysis considers uncertainties for the most important sensitivity parameters
##Output is transformed and plotted in ggplot2.

```{r uncertainty yedoma}
sense_pars_y <- c("undisturbed_yedoma_0", "c_yedoma_0", 
                  "old_lake_yedoma_NEE", 
                  "old_lake_yedoma_0", "a_dynamic_yedoma", 
                  "undisturbed_yedoma_NEE", "alas_yedoma_NEE")
par_ranges_y <- lake_pars[sense_pars_y] %>% 
  enframe() %>% 
  mutate(min = value - 0.4 * value, max = value + 0.4 * value) %>% 
  select(min, max) %>% 
  as.data.frame()
rownames(par_ranges_y) <- sense_pars_y
sR_n <- sensRange(func = NECB_model, parms = lake_pars, start_time=2000, end_time=2300, 
                dist = "grid",
                sensvar = "Total_NECB", parRange = par_ranges_y, 
                num = 3^nrow(par_ranges_y))
# tidy data
yedoma_glb_sens <- sR_n %>% 
  gather(year, necb, -one_of(rownames(par_ranges_y))) %>% 
  mutate(year = str_extract(year, "[0-9]+$") %>% parse_number()) %>% 
  select(year, everything())

# plot
yedoma_glb_sens_summ <- yedoma_glb_sens %>% 
  group_by(year) %>% 
  summarize(necb_mean = mean(necb),
            necb_min = min(necb), necb_max = max(necb),
            necb_sd = sd(necb)) %>% 
  ungroup() %>% 
  mutate(necb_psd = necb_mean + necb_sd, necb_msd = necb_mean - necb_sd)
write_csv(yedoma_glb_sens, "lake_global-sensitivity_yedoma.csv")
write_csv(yedoma_glb_sens_summ, "lake_global-sensitivity_yedoma_summary.csv")
ggplot(yedoma_glb_sens_summ, aes(x = year)) +
  # mean+-sd
  geom_ribbon(aes(ymin = necb_msd, ymax = necb_psd), alpha = 0.2) +
  # mean
  geom_line(aes(y = necb_mean)) +
  labs(title = "Uncertainty in Lake Thermokarst NECB (Yedoma)",
       x = "Year", y = "NECB (PgC)") +
 theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```


```{r uncertainty non yedoma}
sense_pars_n <- c("undisturbed_non_0", "b_non", "old_lake_non_NEE", "old_lake_non_0", "undisturbed_non_CH4", "undisturbed_non_NEE")
par_ranges_n <- lake_pars[sense_pars_n] %>% 
 enframe() %>% 
  mutate(min = value - 0.4 * value, max = value + 0.4 * value) %>% 
  select(min, max) %>% 
  as.data.frame()
rownames(par_ranges_n) <- sense_pars_n
sR_n <- sensRange(func = NECB_model, parms = lake_pars, 
                start_time=2000, end_time=2300, dist = "grid",
                sensvar = "Total_NECB", parRange = par_ranges_n, 
                num = 3^nrow(par_ranges_n))
# tidy data
non_glb_sens <- sR_n %>% 
  gather(year, necb, -one_of(rownames(par_ranges_n))) %>% 
  mutate(year = str_extract(year, "[0-9]+$") %>% parse_number()) %>% 
  select(year, everything())
# plot
non_glb_sens_summ <- non_glb_sens %>% 
  group_by(year) %>% 
  summarize(necb_mean = mean(necb),
            necb_min = min(necb), necb_max = max(necb),
            necb_sd = sd(necb)) %>% 
  ungroup() %>% 
  mutate(necb_psd = necb_mean + necb_sd, necb_msd = necb_mean - necb_sd)
write_csv(yedoma_glb_sens, "lake_global-sensitivity_non-yedoma.csv")
write_csv(non_glb_sens_summ, "lake_global-sensitivity_summary_non-yedoma.csv")
ggplot(non_glb_sens_summ, aes(x = year)) +
  # mean+-sd
  geom_ribbon(aes(ymin = necb_msd, ymax = necb_psd), alpha = 0.2) +
  # mean
  geom_line(aes(y = necb_mean)) +
  labs(title = "Uncertainty in Lake Thermokarst NECB (Non-Yedoma)",
       x = "Year", y = "NECB (PgC)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```

## Combined
##Look at the sensitivity to the full set of yedoma and nonyedoma parameters. This takes a long time to run so we save the results.
```{r uncertainty combined}
# combined parameters
p_unique <- par_ranges_n[!rownames(par_ranges_n) %in% rownames(par_ranges_y), ]
par_ranges_c <- rbind(par_ranges_y, p_unique)
f <- "lake_global-sensitivity.rds"
#if (!file.exists(f)) {
  # sensitivity analysis
  sR_c <- sensRange(func = NECB_model, parms = lake_pars, 
                    start_time=2000, end_time=2300, dist = "grid",
                    sensvar = "Total_NECB", parRange = par_ranges_c, 
                    num = 2^nrow(par_ranges_c))
  saveRDS(sR_c, f)
#}
sR_c <- readRDS(f)
# tidy data
combined_glb_sens <- sR_c %>% 
  gather(year, necb, -one_of(rownames(par_ranges_c))) %>% 
  mutate(year = str_extract(year, "[0-9]+$") %>% parse_number()) %>% 
  select(year, everything())
# plot
glb_sens_summ_combined <- combined_glb_sens %>% 
  group_by(year) %>% 
  summarize(necb_mean = mean(necb),
            necb_min = min(necb), necb_max = max(necb),
            necb_sd = sd(necb)) %>% 
  ungroup() %>% 
  mutate(necb_psd = necb_mean + necb_sd, necb_msd = necb_mean - necb_sd)
write_csv(glb_sens_summ_combined, "lake_global-sensitivity_summary.csv")
ggplot(glb_sens_summ_combined, aes(x = year)) +
  # mean+-sd
  geom_ribbon(aes(ymin = necb_msd, ymax = necb_psd), alpha = 0.2) +
  # mean
  geom_line(aes(y = necb_mean)) +
  labs(title = "Uncertainty in Lake Thermokarst NECB",
       x = "Year", y = "NECB (PgC)") +
  theme(legend.position = "bottom", 
        legend.key=element_rect(fill=NA,colour=NA),
        panel.border=element_blank(),
        axis.line=element_line(),
        panel.background=element_rect(fill="white"),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())
```

# Take CH4 data from output of model and calculate radiative forcing
## #CH_4 and CO_2 started in PgC/km2 terrain and now are converted to gC/m2
## Fluxes need to be summed up across the area of interest (i.e. 900,900 km2) and expressed in kg; positive fluxes are net release to the atmosphere and negative fluxes are net uptake from the atmosphere. 
## Ensure that fluxes are expressed in CO2 and CH4 (1 g C-CO2 = 44/12 g CO2; 1g C-CH4 = 16/12 g CH4.
## Below we are calculating the rate of CH4 and CO2 inputs specific to a year.  This is to allow me to estimate instantaneous RF that takes into the account the current year's perturbation to the atmosphere plus the legacy of previous years.
## length equals 300 yrs
## n represents duration of model run up to 2300 in years
## j is the cohort and i is the year of that cohort's effect
## historic_CH4 starts at zero, effects are summed over time because we are taking one time step and adding in the next year's flux
```{r radiative forcing}
lake_out$CH4g_m2<- lake_out$CH4*10^9/1232000
lake_out$CO2g_m2<-lake_out$CO2*10^9/1232000

lake_out$CH4_kg <- lake_out$CH4g_m2 * -1 * 16/12 * 1232000 * 1000 * 1000 / 1000
lake_out$CO2_kg <- lake_out$CO2g_m2 * -1 * 44/12 * 1232000 * 1000 * 1000 / 1000

fCH4 <- 1.3       
ACH4 <-  1.3e-13     
LifeCH4 <- 12      
n <- length(lake_out$time)

lake_out$delta_CH4_kg[1]<-0
lake_out$delta_CO2_kg[1]<-0
for (t in 2:length(lake_out$time)){
  lake_out$delta_CH4_kg[t] = lake_out$CH4_kg[t]-lake_out$CH4_kg[t-1]
  lake_out$delta_CO2_kg[t] = lake_out$CO2_kg[t]-lake_out$CO2_kg[t-1]
}

# RF METHANE
for (i in 1:n) {
  historic_CH4 <- 0
  for (j in 1:i) {
    historic_CH4 <- historic_CH4 + lake_out$delta_CH4_kg[j] * exp((j - i) / LifeCH4) 
  }
  lake_out$historic_CH4[i] <- historic_CH4
  lake_out$RFCH4[i] <- fCH4 * ACH4 * ((historic_CH4))
}


# RF CARBON DIOXIDE
fCO2 <- 1
ACO2 <- 0.0198e-13   
LifeCO2_1 <- 1e8
LifeCO2_2 <- 421
LifeCO2_3 <- 71
LifeCO2_4 <- 21
LifeCO2_5 <- 3.4

n <- length(lake_out$time)

lake_out$CO2_1 <- lake_out$delta_CO2_kg * 0.176
lake_out$CO2_2 <- lake_out$delta_CO2_kg * 0.138
lake_out$CO2_3 <- lake_out$delta_CO2_kg * 0.186
lake_out$CO2_4 <- lake_out$delta_CO2_kg * 0.242
lake_out$CO2_5 <- lake_out$delta_CO2_kg * 0.259

for (i in 1:n) {
  historic_CO2_1 <- 0
  historic_CO2_2 <- 0
  historic_CO2_3 <- 0
  historic_CO2_4 <- 0
  historic_CO2_5 <- 0

  for (j in 1:i) {
    historic_CO2_1 <- historic_CO2_1 + exp((j - i) / LifeCO2_1) * lake_out$CO2_1[j]
    historic_CO2_2 <- historic_CO2_2 + exp((j - i) / LifeCO2_2) * lake_out$CO2_2[j]
    historic_CO2_3 <- historic_CO2_3 + exp((j - i) / LifeCO2_3) * lake_out$CO2_3[j]
    historic_CO2_4 <- historic_CO2_4 + exp((j - i) / LifeCO2_4) * lake_out$CO2_4[j]
    historic_CO2_5 <- historic_CO2_5 + exp((j - i) / LifeCO2_5) * lake_out$CO2_5[j]
  }
  lake_out$historic_CO2[i] <- historic_CO2_1 + historic_CO2_2 + historic_CO2_3 + historic_CO2_4 + historic_CO2_5
  lake_out$RFCO2[i] <- fCO2 * ACO2 * ((lake_out$historic_CO2[i]))
  
  lake_out$RFC=lake_out$RFCO2 + lake_out$RFCH4
}
write_csv(lake_out, "lake_model.csv")
```

#Dual axis plot.
```{r dual axis plot}
par(mar = c(5.1, 4.1, 2.1, 4.1))
gs <- glb_sens_summ_combined
plot(Total_NECB ~ time, data = lake_out,
     ylim = range(lake_out$Total_NECB), xlim = c(1900, 2300),
     type = "l", lwd = 1.5, lty = 2, col = "black",
     yaxt = "n", xlab = "Year", ylab = NA)
axis(4)
mtext("Cumulative NECB (PgC)", side = 4, line = 2)
# sd bands
gs <- filter(gs, !is.na(year))
polygon(x = c(gs$year, rev(gs$year)),
        y = c(gs$necb_msd, rev(gs$necb_psd)),
        col = "grey70", border = "transparent")
# post 2000
lines(Total_NECB ~ time, data = lake_out %>% filter(time > 2000),
      col = "black", lwd = 1.5)
par(new = TRUE)
plot(RFCO2 ~ time, data = lake_out,
     col = "#e41a1c", type = "l", lwd = 1.5, lty = 2,
     ylim = c(-0.01, 0.1),
     xaxt = "n", yaxt = "n", xlab = NA, ylab = NA)
# post 2000
lines(RFCO2 ~ time, data = lake_out %>% filter(time > 2000), 
      col = "#e41a1c", lwd = 1.5)
axis(2)
mtext(expression("Net Radiative Forcing"~(W ~m^2)), side = 2, line = 2)
lines(RFCH4 ~ time, data = lake_out, col = "#377eb8", lwd = 1.5, lty = 2)
# post 2000
lines(RFCH4 ~ time, data = lake_out %>% filter(time > 2000),
      col = "#377eb8", lwd = 1.5)
legend(x = 1890, y = 0.07, lty = 1, bty = "n",
       col = c("#e41a1c", "#377eb8", "black"), 
       legend = c(expression(RF~CO[2]), expression(RF ~ CH[4]), "NECB (mean±sd)"))
```
